Imminent range shifts in transboundary fish stocks challenge their management under climate change

### 2021-03-22 Scripts/Manuscript.Rmd
Addressed william's comments

Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn]. Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk]). As a result of climate change, many marine species have changed their distributions towards higher latitude, deeper water or have followed local temperature gradients [@Poloczanska:2016kk]. Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds]. The projected risks of impacts can be reduced through improving the effectiveness of current fisheries management [@Gaines:2018sg; @Cheung:2018dh], including the fisheries management of species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].

@@

On average, stocks that see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of the revenue generated by fisheries that target transboundary stocks within their EEZs between 2004 and 2010. However, large variation exists in the contribution of stocks with emerging transboundary index to the current revenue participation across EEZs. In some cases, like El Salvador and New Zealand, stocks with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries in their respective EEZs between 2004 and 2014. On the contrary, stocks with emerging transboundary index from the Marshall Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A). In addition, some EEZs are projected to have a few emerging stocks that represent a large proportion of revenue from transboundary stocks like Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively). Overall, Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks to fishing revenue is less than 20%. This indicates that, despite having a large proportion of transboundary stocks emerging, the economic contribution of these are relatively small to the overall revenue generated by fisheries targetting transboundary stocks in these countries.

@@

I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods – Equation \@ref(eq:five)). By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A). The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)). However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high. For example, Guatemala shares *Panulirus gracilis* stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century. Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)). My results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A). These EEZs will changes in the stock share ratio beyond the threat point in one third (33% $\pm$ 21%) of their transboundary stocks (Figure \@ref(fig:toeFigure5)A). Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin are projected to have changes beyond the threat point for all of the transboundary stocks analyzed, while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) are the only countries with less than 10% of stocks changing beyond the threat point.

@@

In this chapter, I investigate the timing and intensity of changes in the sharing of transboundary stocks between countries. My findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world. Such early emergence of range shifts indicate the urgency of addressing the challenges that climate change is posing on international fisheries management. Moreover, I identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management. Finally, I show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to levels below of that needed to encourage cooperative resource management between countries. Adaptation measures are therefore needed to reduce the impacts of species range shifts on current international transboundary fisheries treaties.

>>>>

Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, including warming, acidification and reducing oxygen content [@IPCC:2019tn]. Distributions of marine species are closely related to the environment and the species’ preferences to environmental conditions (e.g., temperature, oxygen, salinity) [@Hutchinson:1957nt; @Nelson:2016kk]. As a result of climate change, many marine species have changed their distributions towards higher latitude, deeper water or have followed local temperature gradients [@Poloczanska:2016kk]. Biogeography of marine species is projected to continue to shift as ocean conditions are changing in the $21^{st}$ century under climate change [@Cheung:2010dt], impacting fisheries production and compromising our capacity to reach international sustainability goals such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Barange:2014kr; @Singh:2017ds]. The projected risks of impacts can be reduced through improving the effectiveness of current fisheries management [@Gaines:2018sg; @Cheung:2018dh], including the fisheries management of species that cross international borders, i.e., shared stocks [@Gaines:2018sg; @Pinsky:2018cb].

@@

On average, stocks that see their transboundary index emerge by 2100 represent 27% $\pm$ 23% of the revenue generated by fisheries that target transboundary stocks within their EEZs between 2004 and 2010. However, large variation exists in the contribution of stocks with emerging transboundary index to the current revenue participation across EEZs. In some cases, like El Salvador and New Zealand, stocks with emerging transboundary index represent less than 1% of the revenue generated from transboundary fisheries in their respective EEZs between 2004 and 2014. On the contrary, stocks with emerging transboundary index from the Marshall Islands and Tokelau represent over 90% of the revenue generated from fisheries targeting transboundary stocks within their respective EEZs during the same time period (Figure \@ref(fig:toeFigure2)A). In addition, some EEZs are projected to have a few emerging stocks that represent a large proportion of revenue from transboundary stocks like Peru and Finland (with 10 and 5 emerging stocks representing 86% and 78% of average fishing revenue between 2004 to 2010, respectively). Overall, Morocco and Malaysia are the only two countries amongst the top 10 countries with the larger number of stocks with emerging transboundary index (n = 56 and 53, respectively) where the contribution of such stocks to fishing revenue is less than 20%. This indicates that, despite having a large proportion of transboundary stocks emerging, the economic contribution of these stocks are relatively small to the overall revenue generated by fisheries targeting transboundary stocks in these countries.

@@

I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods – Equation \@ref(eq:five)). By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A). The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)). However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high. For example, Guatemala shares *Panulirus gracilis* stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century. Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)). My results suggest that 87% (n = 238) of the world EEZs will experience changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A). These EEZs are projected to have changes in the stock share ratio beyond the threat point in one third (33% $\pm$ 21%) of their transboundary stocks (Figure \@ref(fig:toeFigure5)A). Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin are projected to have changes beyond the threat point for all of the transboundary stocks analyzed, while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) are the only countries with less than 10% of stocks changing beyond the threat point.

@@

In this chapter, I investigate the timing and intensity of changes in the sharing of transboundary stocks between countries. My findings highlight the early emergence of transboundary distribution shifts from historical variations for most of the shared stocks in the world. Such early emergence of range shifts indicates the urgency of addressing the challenges that climate change is posing on international fisheries management. Moreover, I identify regional patterns of high climate risk to important fish stocks and associated fisheries at the present-day that will require immediate responses from transboundary fisheries management. Finally, I show that in many instances around the world, countries’ fisheries will experience changes in the sharing of many transboundary stocks to levels below of that needed to encourage cooperative resource management between countries. Adaptation measures are therefore needed to reduce the impacts of species range shifts on current international transboundary fisheries treaties.
